---
title: "In Class DT"
author: "Brian Wright"
date: "December 7, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Congrats! You just graduated UVA's MSDS program and got a job working at the 
Treasury Department. In partnership with Congress members the Treasury has been 
ask to come up with innovative ways to create tax policy. In doing so they 
want to be able to reliably predict whether American workers make more than 
$50,000 a year and also which variables seem to be most contributing 
to predicting this outcome. 

You would like to be able to explain the model to the mere mortals 
around you but need a fairly robust and flexible approach so you've 
chosen to use decision trees to get started and will possibly move 
to a ensemble model if needed. 

In doing so, similar to  great data scientists of the past 
you remembered the excellent education provided 
to you at UVA in a undergrad data science course and have outline 
20ish steps that will need to be undertaken to complete this task 
(you can add more or combine if needed).  As always, you will need 
to make sure to #comment your work heavily. 


```{r}
#Data is combination of factor level variables and contiuous
#Ask yourself which variables will work and won't so do data cleaning
#Make sure variables are labeled correctly (factors or contiuous)
#You don't need to normalize continuous variables
#Have to make <= a factor variable, like 0's and 1's

#Lots of things should be factor levels, not characters

```

 Footnotes: 
-	You can add or combine steps if needed
-	Also, remember to try several methods during evaluation and always be 
mindful of how the model will be used in practice.
- Make sure all your variables are the correct type (factor, character, etc.


```{r}
#install.packages("rio")
library(rio)
library(plyr)
library(tidyverse)
library(rpart)
library(psych)
library(pROC)
#install.packages("rpart.plot")
library(rpart.plot)
install.packages("rattle")
library(rattle)
library(caret)
library(C50)
install.packages("mlbench")
library(mlbench)

library(mltools)
library(data.table)
library(caret)
```

```{r}
#1 Load the data, check for missing data and ensure the labels are correct.

url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data"
xx <- readr::read_csv(url,col_names= FALSE)
view(xx)
```

```{r}
# Output Dimensions
cat("Shape:", dim(xx), "\n")
```

```{r}
# Peep first five rows
head(xx)
```

```{r}
# Copy the Data Set
DATA = xx
```

```{r}
# Improve Quality of Column Labels
names(DATA) <- c(
  'age', 'workclass', 'fnlwgt', 'education', 'education_num', 'marital_status', 'occupation', 'relationship',
  'race', 'sex', 'capital_gain', 'capital_loss', 'hours_per_week', 'native_country', 'income'
)
head(DATA)
```

```{r}
cat("Columns:", names(DATA))
```

```{r}
# Remove Null and Fake Null ("?") Data Points
cat("NULL:", sum(is.na(DATA)))
DATA <- DATA %>% 
  filter(workclass != "?", occupation != "?", native_country != "?")
cat("New Shape:", dim(DATA))
```

### Step 2
Ensure all the variables are classified correctly including the target variable
```{r}
# View Data Set structure
str(DATA)
```

```{r}
# Convert Categorical to Factor Variables
DATA[,c(2, 5:9, 13, 14)] <- lapply(DATA[,c(2, 5:9, 13, 14)], as.factor)
str(DATA)
```

```{r}
# Explore Categorical Variables
table(DATA$`workclass`) # Remove ? 
table(DATA$`education`) # Factor Collapse
table(DATA$`education_num`) # DELETE? -- Duplicate
table(DATA$`marital_status`) # Factor Collapse?
table(DATA$`occupation`) # Remove ? 
table(DATA$`relationship`) # Fine?
table(DATA$`race`) # Fine?
table(DATA$`sex`) # Fine?
table(DATA$`native_country`) # Remove ? 
table(DATA$`income`) # Fine
```

```{r}
# ----- Factor Collapse Multiple Variables -----
# education
DATA$`education` <- fct_collapse(
  DATA$`education`,
  Pre_HS=c("Preschool", "1st-4th", "5th-6th", "7th-8th"),
  HS=c("9th", "10th", "11th", "12th", "HS-grad"),
  College=c("Some-college", "Assoc-voc", "Assoc-acdm", "Bachelors"),
  Post_Grad=c("Masters", "Prof-school", "Doctorate")
)

# Delete duplicate education_num
DATA <- DATA[, -5]
cat("New Shape:", dim(DATA))

# marital_status
DATA$`marital_status` <- fct_collapse(
  DATA$`marital_status`,
  Never_Married= "Never_married",
  Married=c("Married-civ-spouse", "Married-spouse-absent", "Married-AF-spouse"),
  Separated=c("Separated", "Divorced", "Widowed"),
)
table(DATA$marital_status)

# native_country
DATA$`native_country` <- fct_collapse(
  DATA$`native_country`,
  North_America=c("Canada", "North_America", "United-States"),
  South_America=c("Laos", "Nicaragua", "Puerto-Rico", "Trinadad&Tobago", "South", 
                  "Peru","Outlying-US(Guam-USVI-etc)", "Mexico", "Jamaica", 
                  "Honduras", "Haiti", "Guatemala", "El-Salvador", "Ecuador", 
                  "Dominican-Republic", "Cuba", "Columbia"),
  Europe=c("Germany", "France", "England", "Greece", "Italy", "Hungary", 
           "Ireland", "Portugal", "Poland", "Scotland", "Yugoslavia", "Scotland", 
           "Iran", "Holand-Netherlands"),
  Asia=c("Asia", "China", "Thailand", "Vietnam", "Taiwan", "Philippines", "Japan",
         "India", "Hong", "Cambodia")
)
table(DATA$native_country)
# TARGET: income

# Convert Categorical to Factor Variables
DATA[,c(2, 5:9, 13, 14)] <- lapply(DATA[,c(2, 5:9, 13, 14)], as.factor)
str(DATA)

table(DATA$income)
#Collapse Target Variable
DATA$income <- fct_collapse(DATA$income,
                                      rich=">50K",
                                      poor="<=50K")
str(DATA)
```

```{r}
#3 Don't check for correlated variables....because it doesn't matter with 
# Decision Trees...the make local greedy decisions. 
```

```{r}
#4 Guess what, you also don't need to standardize the data, 
#because DTs don't give a ish, they make local decisions...keeps getting easier
```

```{r}
#5 Determine the baserate or prevalence for the classifier, 
# what does this number mean?  

# View # of Data Points by Class
table(DATA$`income`)
less_than = 22654 # <50k
more_than = 7508  # >50k
# Calculate Prevalence
  # ~62% of people in the dataset make less than $50,000
prevalence = less_than / (less_than + more_than)
prevalence
```

```{r}
#6 Split your data into test, tune, and train. (70/15/15)

part_index_1 <- caret::createDataPartition(DATA$income,
                                           times=1,
                                           p = 0.70,
                                           groups=1,
                                           list=FALSE)

train <- DATA[part_index_1, ]
tune_and_test <- DATA[-part_index_1, ]

#The we need to use the function again to create the tuning set 

tune_and_test_index <- createDataPartition(tune_and_test$income,
                                           p = .5,
                                           list = FALSE,
                                           times = 1)

tune <- tune_and_test[tune_and_test_index, ]
test <- tune_and_test[-tune_and_test_index, ]

dim(train)
dim(tune)
dim(test)
```

```{r}
#7 Build your model using the training data and default settings in caret, 
# double check to make sure you are using a cross-validation training approach

# Choose the features and classes
#TARGET = Quality
#Features are what we use to learn about the target 


features <- train[,-14]
target <- train$income


str(features)
str(target)

#Cross validation process
fitControl <- trainControl(method = "repeatedcv",
                          number = 10,
                          repeats = 5, 
                          returnResamp="all",
                          classProbs = TRUE,
                          allowParallel = TRUE)

grid <- expand.grid(.winnow = c(TRUE,FALSE), 
                    .trials=c(1,5,10,15,20), 
                    .model="tree")
set.seed(1984)
DATA_mdl <- train(x=features,
                y=target,
                method="C5.0",
                tuneGrid=grid,
                trControl=fitControl,
                verbose=TRUE)
DATA_mdl
```

```{r}
#8 View the results, what is the most important variable for the tree? 
```

```{r}
#9 Plot the output of the model to see the tree visually 
```

```{r}
#10 Use the validation set and the predict function with your model to the
# estimate the target variable.
```

```{r}
#11 Compare the predicted values to those of the actual by generating a 
# matrix ("by-hand").
```

```{r}
#12 Use the the confusion matrix function to check a variety of metrics 
# and comment on the metric that might be best for this type of analysis given your question.  
```

```{r}
#13 Generate a ROC and AUC output, interpret the results
```

```{r}
#14 Use the predict function to generate percentages, then select several 
# different threshold levels using the confusion matrix function and 
# interpret the results. What patterns did you notice, did the evaluation metrics change? 
```

```{r}
#TUNE PHASE
#15 Based on your understanding of the model and data adjust several of the hyper-parameters via the built in train control function in caret or build and try new features, does the model quality improve? If so how and why, if not, why not?
# Use this link: https://rdrr.io/cran/caret/man/trainControl.html to select changes,
# you aren't expected to understand all these options but explore one or two and 
# see what happens. 

```

```{r}
#16 Once you are confident that your model is not improving, via changes 
# implemented on the training set and evaluated on the the validation set (item 16), predict with the test set and report a final evaluation of the model. Discuss the output in comparison with the previous evaluations.  
```

```{r}
#17 Summarize what you learned along the way and make recommendations on how this could be used moving forward, being careful not to over promise. 
```

```{r}
#18 What was the most interesting or hardest part of this process and what questions do you still have? 
```


